---
title: "PSY6422 Mini Project"
format: html
editor: visual
---

# 1.0 Overview

\[Provide overview of what this page contains\]

## 1.1 Packages

```{r}
#| message: FALSE

# Install packages - update as you progress through the project

# Load packages
library(tidyverse)
library(here)
```

# 2.0 Data Origins

The data used in this project was found in a Crowdsourced list of ['Open Psychological Datasets'](https://docs.google.com/spreadsheets/d/1ejOJTNTL5ApCuGTUciV0REEEAqvhI2Rd2FCoj7afops/edit?gid=0#gid=0https://docs.google.com/spreadsheets/d/1ejOJTNTL5ApCuGTUciV0REEEAqvhI2Rd2FCoj7afops/edit?gid=0#gid=0). In this list was the [OpenNeuro Dataset of Metadata](https://docs.google.com/spreadsheets/d/1rsVlKg0vBzkx7XUGK4joky9cM8umtkQRpJ2Y-5d6x7c/edit?gid=762232233#gid=762232233). OpenNeuro is a data archive that shares public datasets for others to use in their research. The meta-dataset used in this project consists of a list of neuroimaging studies that have used a range of imaging modalities. This data was collected by researchers uploading their data to the data archive. The dataset was downloaded on the 10th of October 2025 and therefore may not contain all data that is in the live metadata online.

## 2.1 Load Data

```{r}
#| message: FALSE
#| results: 'hide'

# Load Open Neuro Metadata dataset
neuroscience_raw <- read_csv(here("datasets", "neuroscience_metadata.csv"))
```

## 2.2 Raw Data Summary and Variables

```{r}
#| message: FALSE

# Show the first ten rows of the raw data
head(neuroscience_raw, 10)

# Provide a summary of the dataset
summary(neuroscience_raw)
```

As seen from viewing the first ten rows of the raw data, there are 48 columns. Due to the this, not all variables are shown when using the head() function therefore, I have used the summary() function to gain information about the hidden variables. Using this function determines that first 23 of the 48 columns are variables with the remaining, 25 columns containing NAs. From these 23 variables, 3 will be retained for the analysis conducted in this project. These variables are as follows:

| Variable Name | Meaning                                       |
|---------------|-----------------------------------------------|
| made_public   | the date the dataset was made public          |
| modalities    | the imaging modalities used in the study      |
| species       | the species of the subjects used in the study |

# 3.0 Research Question

\[In plain English, a simple statement of what question(s) your visualisation will attempt to address\]

The aim of this project is to visualise how has use of neuroimaging techniques in human studies changed over time. The visualisation produced to answer this research question will demonstrate the count of individual **(provide examples)** and multimodal imaging techniques over time from **\[beginning year\]** to **\[end year\]**.

# 4.0 Data Preparation

\[Steps taken to clean the data, exclude outliers, sanity check the data range and values, create summary statistics, grouped variables, etc. Show the first few rows of the processed data, if possible. Showing the code which does this where relevant\]

## 4.1 Data Cleaning

Firstly the variables and empty columns were removed from the dataset so that only the variables needed for the visualisation remained. In addition to this, two new columns were created: 'ID' and 'date_published'. The 'ID' column makes the dataset easier to read while the 'date_published' column converts the 'made_public' data from a character variable to a date variable. The 'made_public' column was then removed to avoid confusion with the new 'date_published' column.

```{r}
## Clean data -----------------------------------------------
### Make dataset with just the variables needed for the -
### project
neuroscience_data <- neuroscience_raw %>% 
  select(made_public, modalities, species) %>% # Only retain these columns
  drop_na() %>% # Remove all rows containing NAs
  mutate(ID = row_number(), .before = 1, # Create an ID column to the left of the dataset
         date_published = as.Date(made_public, "%m/%d/%Y")) %>% # Create a date_published column with a date data type
  select(ID, date_published, modalities, species) # Remove the made_public column
```

Next the 'species' column was filtered to only include human studies as the research question involves investigating the neuroimaging techniques in humans.

```{r}
### Remove subjects that aren't human
human_data <- neuroscience_data %>% 
  filter(species == "Human")

### Sanity Check: check that there aren't any non-human species remaining
for(i in human_data$species){
  if(i != "Human"){
    print("Test Failed") # If test is passed nothing should be printed
  }
}
```

## 4.2 Renaming 'modalities' values

The values in the modalities column were not suitable for this project as they were inconsistent strings. In order to make the data ready for producing a visualisation, the values were recoded into 6 categories: 'multimodal', 'eeg', 'meg', 'dMRI', 'fMRI', and 'sMRI'.

```{r}
### View the number of modalities in the dataset and save as a dataset
modality_count <- human_data %>% 
  group_by(modalities) %>% 
  count()

### Remove unspecified groups in modalities variable and rename with imaging categories
modality_data <- human_data %>% 
  filter(modalities != "beh") %>% 
  mutate(modalities = case_match(modalities, 
                                 c("bold, events, t1w",
                                   "eeg, nirs",
                                   "mri_diffusion, mri_functional, mri",
                                   "mri_diffusion, mri_functional, mri_structural, eeg, mri",
                                   "mri_diffusion, mri_functional, mri_structural, mri",
                                   "mri_diffusion, mri_structural, eeg, mri",
                                   "mri_diffusion, mri_structural, mri",
                                   "mri_diffusion, mri_structural, mri_functional, mri",
                                   "mri_diffusion, mri_structural, mri_functional, mri, pet",
                                   "mri_diffusion, mri_structural, mri_functional, mri_perfusion, mri",
                                   "mri_functional, mri, eeg",
                                   "mri_functional, mri_diffusion, mri_structural, eeg, mri, beh",
                                   "mri_functional, mri_diffusion, mri_structural, meg, mri",
                                   "mri_functional, mri_diffusion, mri_structural, mri",
                                   "mri_functional, mri_perfusion, mri_structural, mri",
                                   "mri_functional, mri_structural, eeg, mri",
                                   "mri_functional, mri_structural, mri",
                                   "mri_functional, mri_structural, mri, beh",
                                   "mri_functional, mri_structural, mri, eeg",
                                   "mri_functional, mri_structural, mri, eeg, beh",
                                   "mri_functional, mri_structural, mri, ieeg",
                                   "mri_functional, mri_structural, mri_diffusion, mri",
                                   "mri_functional, mri_structural, pet_dynamic, mri, pet",
                                   "mri_functional, pet_static, mri_structural, pet_dynamic, mri, pet",
                                   "mri_structural, eeg, mri",
                                   "mri_structural, ieeg, mri",
                                   "mri_structural, meg, mri",
                                   "mri_structural, meg, mri, beh",
                                   "mri_structural, mri, pet",
                                   "mri_structural, mri_diffusion, mri",
                                   "mri_structural, mri_diffusion, mri_functional, mri",
                                   "mri_structural, mri_diffusion, mri_functional, mri, eeg",
                                   "mri_structural, mri_functional, ieeg, mri",
                                   "mri_structural, mri_functional, mri",
                                   "mri_structural, mri_functional, mri, eeg",
                                   "mri_structural, pet, mri",
                                   "pet_dynamic, mri_functional, mri_structural, mri, pet",
                                   "t1w, bold, events",
                                   "t1w, bold, events, fieldmap",
                                   "t1w, channels, eeg, events, bold") ~ "multimodal",
                                 c("channels, eeg, electrodes, events", 
                                   "channels, eeg, events",
                                   "eeg, beh",
                                   "ieeg",
                                   "ieeg, eeg") ~ "eeg",
                                 "meg, beh" ~ "meg",
                                 "mri_diffusion, mri" ~ "dMRI",
                                 "mri_functional, mri" ~ "fMRI",
                                 c("mri_structural, mri",
                                   "t1w") ~ "sMRI",
                                 .default = modalities))

### Sanity check: Ensure the total number of observations is 846 in the human_data 
### and that the only categories are "multimodal", "eeg", "meg", "dMRI", "fMRI", and "sMRI"
modality_count <- modality_data %>% 
  group_by(modalities) %>% 
  count() 

sum(modality_count$n)
```

After completion of this step, the data was ready for visualisation. Below is a preview of the processed data.

```{r}
### If the above sanity check is passed then save the data as below
neuroimaging_overtime <- modality_data

### Preview the final dataset
head(neuroimaging_overtime, 10)
```

# 5.0 Visualisation

\[Graph or graphs. Documentation explaining any motivation (although good graph labelling is better than explanation in the accompanying text). Code for producing graph(s)\]

# 6.0 Summary

\[Brief thoughts on what you have learnt, what you might do next if you had more time / more data\]
